---
title: "How to use hydramse"
output: rmarkdown::html_vignette

vignette: >
  %\VignetteIndexEntry{How to use hydramse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The functions in this package allow you to perform a sequence of events resulting in an MSE (Management Strategy Evaluation) type of workflow (Assumes package hydradata and hydra_sim (tpl file)). This is not a stand alone package and needs to be used in conjunction with the aforementioned packages

1. Simulate a set of parameters that are assumed somewhat unknown (assumes hydradata)
1. Assess the validity of this parameter set.
1. If assumed valid ,define a set of scenarios with ranges of exploitation rates (otherwise return to step 1)
1. Create dat and pin files for each scenario. These are input files for Hydra model (assumes hydradata)
1. Run Hydra a number of times (with stochasticity) (assumes hydra_sim)
1. Process the output and produce plots
1. Repeat until n valid sets of parameters have been obtained
1. Compare all output (NOT YET DONE)

## Model Setup

Decisions need to be made regarding how long to run the model for in the absence of fishing (the number of years). This decision is based on how long the model takes to reach some kind of equilibrium (if it does at all). An equilibrium may not be reached if a species stock recruitment relationship is highly overcompensatory. This may result in a highly cyclical behavior, possiblly around an average. It is recommended that the user specify the total length of the model run.

Once specified, the underlying data used to create Hydra's data files (dat and pin) needs to be extended to accomodate these additional years of data.

``` {r echo=TRUE,eval = FALSE}
newData <- hydramse::lengthen_hydra_data(hydradata::hydraDataList,nYrs=150)
```

Set up the locations for output (create if necessary):
``` {r  echo=TRUE,eval =FALSE}
outDirForDatPin <- here::here("darwin") # folder for dumping hydra output prior to analysis
outPath <- here::here("successfulSims") # folder containing parameters of successful model runs

if(!dir.exists(outDirForDatPin)){dir.create(outDirForDatPin)}
if(!dir.exists(outPath)){dir.create(outPath)}
```

Also need to set the name of the hydra's exe file
```{r, eval=F,echo=T}
hydraVersion <- "hydra_sim"
```

## Simulation (Darwinian) study

Methods used in the simulation study can be found [here](simulation_documentation.html). The code below should be included in a loop to simulate multiple parameter sets

``` {r  eval=F, echo=T}
  ipass <- 0
  simulationRules <- hydramse::darwinRules
  darwinD <- hydramse::darwinData

  # simulate a set of parameters
  simulatedValues <- hydramse::simulate_parameters(darwinD,simulationRules,SRFunctionChoice = NULL)
  # combine with extended time series data for darwinian use
  hydraD <- hydramse::update_hydra_data(hydradata::hydraDataList,simulatedValues) # update with simulated values
  hydraD <- hydramse::update_hydra_data(hydradata::hydraDataList,noFishingData) # update with simulated values
  # create dat and pin files
  inputOptions <- hydradata::setup_default_inputs(outDir=outDirForDatPin) # gets options for hydraRuns
  inputOptions$scenarioFlag <- "darwin"
  hydraD$flagMSE <- 2 # reduces output from hydra run
  hydraD$recStochastic <- hydraD$recStochastic*0 # no stochasticity used in model run
  hydraD <- hydradata::create_datpin_files(inputOptions,hydraD) # creates dat and pin file
  
  # run hydra 
  #specify the path the input files and the Hydra model executable
  datPath <- paste0(outDirForDatPin,"/",inputOptions$outputFilename,".dat")
  pinPath <- paste0(outDirForDatPin,"/",inputOptions$outputFilename,".pin")
  exePath <- paste(pathToTPL,hydraVersion,sep="/")
  hydramse::run_single_hydra(iseed=1,exePath=exePath,datPath=datPath,pinPath=pinPath)

  # move output file
  if (Sys.info()['sysname']=="Windows") {
    shell(paste0("move *.text ",paste0(outDirForDatPin,"/")),intern=T) # move files to darwin folder
  } else if (Sys.info()['sysname']=="Linux") {
    system(paste0("mv *.text ",paste0(outDirForDatPin,"/")),intern=T) # move files to darwin folder
  }
```

## Assess the validity of the parameter set

Once the mode has run, the output needs to be processed and compared to historic data. If tests are passes this realization is saved for use in an MSE

```{r eval= F,echo=T}
  # processes output from model (1 file)
  output <- hydramse::process_darwin_output(outDirForDatPin)
  biomass <- output$biomass
  catch <- output$catch
  # Do these simulations satisfy the rules(historical equivalence)
  # No Fishing Biomass Rule. assumes equilibrium is reached
  rule1 <- hydramse::rule1_biomass(biomass,nYrsFishing,stockRecruitData$historicBounds,simulationRules)
  if(rule1$pass == F) { next}
  #  Fishing Biomass Rule. use mean of last 10 years
  rule2 <- hydramse::rule2_biomass(biomass,nYrs,stockRecruitData$historicBounds,simulationRules)
  if(rule2$pass == F) { next}
  # Catch Rule
  rule3 <- hydramse::rule3_landings(catch,nYrs,stockRecruitData$historicBounds,simulationRules)
  if(rule3$pass == F) { next}
  
  ipass <- ipass + 1
  # save all data to RDA file for use in MSE
  darwinRes <- list(hydraDarwin=hydraD,simulatedValues=simulatedValues,simulationRules=simulationRules,inputOptions=inputOptions)
  saveRDS(darwinRes,file=paste0(outPath,"/success",ipass,".RDS"))
  
  
```

## Define Scenarios for MSE

Each parameter set that passes the tests are then used in an MSE type study. Details of which can be found in a the publication (Beet & Fogarty, 2020). The parameter set can be read from .Rda file or passed straight from the darwin study. Here we assume it is read from the rda file. This simulated data set is used to create all o fthe dat an dpin files required for each scenario/ exploitation rate combination

```{r,eval=F,echo=T}
rootFolder <- "define_the_folder_where_output_for_all_scenario_runs_be_stored"
darwinRes <- readRDS(file=paste0(outPath,"/success1.RDS")) # read in the RDS
dataToUse <- hydramse::update_hydra_data(hydradata::hydraDataList,darwinRes$simulatedValues)# sets the simulated values to the hydra model data
hcrTypes=c("Fixed","Ramp") # harvest control rules types
hcrLevels=c("Complex","HTSpecies","LTSpecies") #harvest control rule levels
scenarios <- apply(expand.grid(hcrTypes,hcrLevels),1,paste,collapse="")
# Create harvest control rule levels
dataToUse <- hydradata::set_hcr(dataToUse,minMaxExploitations = c(.05,0.4),increment=0.05)
dataToUse$flagMSE <- 1
exploitationRates = round(100*dataToUse$exploitationOptions[dataToUse$Nthresholds,])
# create folder structure for run (given exploitation rates). This creates folders in users working directory
folderStructure <- hydramse::create_folder_setup(rootFolder,exploitationRates = exploitationRates)
scenarios <- unique(apply(as.matrix(subset(folderStructure,select=c("hcrType","hcrLevels"))),1,function(x) paste0(x,collapse = "")))
folderDirs <- apply(folderStructure,1,function(x) paste0(x,collapse = "")) # vector of dirs
outputScenarioDirs <-  here::here(rootFolder,paste0("Exploitation",folderDirs))

for (iscenario in 1:length(outputScenarioDirs)) {
  outputScenarioDir <- outputScenarioDirs[iscenario]
  print(outputScenarioDir)
  # set up scenario types to run
  if (grepl("LT",folderStructure$hcrLevels[iscenario])) speciesFlag <- "low"
  if (grepl("HT",folderStructure$hcrLevels[iscenario])) speciesFlag <- "high"
  if (grepl("omplex",folderStructure$hcrLevels[iscenario])) speciesFlag <- "none"
  scenarioInputs <- hydradata::setup_default_inputs(outDir=outputScenarioDir,scenarioFlag="assessment",temperatureFlag="mean",
                                                    scenarioType=folderStructure$hcrType[iscenario],maxExploitationRate=as.numeric(folderStructure$exploitationRate[iscenario]),
                                                    assessmentSpeciesFlag=speciesFlag)

  # create dat and pin files for scenario
  dataToPrint <- hydradata::create_datpin_files(listOfParameters=scenarioInputs, dataList=dataToUse)
}
```

## Run the hydra model for each scenario

Once the folder structure has been created and the dat and pin files saved into each scenarios respective folder then the model can be run any nimber of times for each scenario. The model can be run in series or in parallel. This is dependent on the number of cores available on your machine or server

``` {r eval=F, echo = T}
   #nCores <- parallel::detectCores()-1
   #cl <- parallel::makeCluster(nCores)

   nSims <- 100 # set number of simulation for each scenario
   exePath <- paste(pathToTPL,hydraVersion,sep="/")
  for (iscenario in 1:dim(folderStructure)[1]) {
    datpinPath <- here::here(rootFolder,paste0("Exploitation",paste0(folderStructure[iscenario,],collapse="")))
    datPath <- paste0(datpinPath,"/",hydraVersion,".dat")
    pinPath <- paste0(datpinPath,"/",hydraVersion,".pin")
    #parallel::parLapply(cl,1:nSims,hydramse::run_single_hydra,exePath=exePath,datPath=datPath,pinPath=pinPath)
    lapply(1:nSims,hydramse::run_single_hydra,exePath=exePath,datPath=datPath,pinPath=pinPath)
    # move files
    if (Sys.info()['sysname']=="Windows") {
      shell(paste0("move *.txt ",paste0(datpinPath,"/indices")),intern=T) # move files to assessment folder
    } else if (Sys.info()['sysname']=="Linux") {
      system(paste0("mv *.txt ",paste0(datpinPath,"/indices")),intern=T) # move files to assessment folder
    }
  }
   #parallel::stopCluster(cl)

```

## Process the output 

nSim files are created for each scenario run and are stored in the "indices" folder under each scenario. These files are processed for all scenarios and exported to RDS files for plotting. The output files from Hydra contain only the variables listed in the variable "indices" below. This speeds up run time. If indices from outside the model are to be included then they need to be read in and passed to process_model_runs()

```{r, eval = F,echo = T}
indices <- c("index_LFI_Catch","index_LFI_Biomass","index_stdev_catch","avByr","est_catch_biomass","index_status_species","index_status_guild")
rD <- read.csv("revenuePricePerPound2012.csv",header=TRUE)
otherData <- list()
otherData$name = rD[,1]
otherData$ppp = rD[,2]
hydramse::process_model_runs(dataToUse,indices,scenarios,rootFolder,outputScenarioDirs,otherData,outputType="indices")

```

## Plotting the output

For each of the harvest control rule type/level combinations:
1. A series of boxplots are made to compare biomass, catch, and species/guild status against exploitation rate.
1. Equivalent Line plots are also made
1. Production function plots are also made for each guild
1. Radar plots are made to compare the performace of each oth the indices

Still in Development
``` {r, eval=F, echo=T}
 hydramse::plot_box_whiskers(here::here(),rootFolder,inputFile="species_bio_rate.rds")
```
